---
title: "K-Means-Example"
author: ""
date: "10/01/2024"
output: pdf_document
---

Set the work directory. 
This MUST be done for every submission
This will determine the default location of all of your data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#This sets the working directory
knitr::opts_knit$set(root.dir = "C:/Users/yunqi/OneDrive - Conestoga College/Conestoga_College/DataScience_Conestoga_2309/Lec_W6")

```

This section is for the basic set up.
It will clear all the plots, the console and the workspace.
It also sets the overall format for numbers.

```{r}
if(!is.null(dev.list())) dev.off()
cat("\014") 
rm(list=ls())
options(scipen=9)
setwd("C:/Users/yunqi/OneDrive - Conestoga College/Course_teach/DataScience8435_24F/Week_6")

```
This section loads and attaches all the necessary packages.

```{r}
#Load packages
#No Packages are needed
```

This section reads in the datafile file

```{r}

DenverC <- read.table("DenverCrimeClstr.txt", sep=",", header=TRUE)

DenverC <- as.data.frame(DenverC)
head(DenverC)

str(DenverC)

```
Transform Data

```{r}

#Convert to a Factor

DenverC$Counsellor <- as.factor(DenverC$Counsellor)

str(DenverC)   #Check Results


```

Set Some Functions
```{r}

#Create a quick standardization function
norm01 <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

#A different standardization function
normn <- function(x) {
  return ((x-mean(x))/sd(x))
}

```


Explore the Data
```{r}


par(mfrow=c(2,2))

for (i in 1:ncol(DenverC)) {
  if (is.numeric(DenverC[,i])) {
      boxplot(DenverC[i], main=names(DenverC)[i],
              horizontal=TRUE, pch=10)
  }
}

par(mfrow=c(1,1))


```
Standardize the Data
Note - Here we are just standardizing the centroids (Population and Crime)

```{r}

DenverC$Pop_Normx <- normn(DenverC$Pop)

DenverC$Crime_Normx <- normn(DenverC$Crime)

head(DenverC)
```


Set Up for Clusters
```{r}

# Creating Variable for Elbow Chart
# Trying for 2 to 8 Clusters
maxk <- 8
nk <- c(2:maxk)
wss <- rep(0,maxk-1)

```
Create Clusters
```{r}

#Set Number of Clusters
k=2

ClstrDen <- kmeans(DenverC[,c(11:12)], iter.max=10, centers=k, nstart=10)
ClstrDen

DenverC$cluster <- factor(ClstrDen$cluster)   # Adding Cluster tags to variables
head(DenverC)

centers <- data.frame(cluster=factor(1:k), ClstrDen$centers)


wss[k-1] <- ClstrDen$tot.withinss
```


Plot 'Elbow'
```{r}

##NOTE - Do NOT use other built in functions to make this chart
##They yield different answers

plot(2:maxk, wss,
     type="b", pch = 19, frame = FALSE,
     main="Elbow Chart for Clusters",
     xlab="Values of K",
     ylab="Total Within-Clusters SS",
     ylim=c(0,max(wss)))
```

```{r}
k=4
ClstrDen <- kmeans(DenverC[,c(11:12)], iter.max=10, centers=k, nstart=10)
ClstrDen

DenverC$cluster <- factor(ClstrDen$cluster)   # Adding Cluster tags to variables
head(DenverC)

centers <- data.frame(cluster=factor(1:k), ClstrDen$centers)
```



Plot the clusters
```{r}
plot(DenverC$Pop_Normx, DenverC$Crime_Normx,
     col=DenverC$cluster, pch=as.numeric(DenverC$cluster))

```

Summarize the Clusters
```{r}
SummClusters <- aggregate(cbind(Pop, Child, Lunch, Income, Crime) ~ cluster,
                          DenverC, FUN=function(x) round(mean(x), 0))
SummClusters

```


